Multilocus Genotype And Mycelial Compatability Group Assessment

### 2017-07-27 doc/RMD/mlg-mcg.Rmd
add code to further examine mlg/mcg relationship

## MLG table

@@

## making a graph

@@

```{r community_counting, fig.width = 5, fig.height = 3}
in_reach <- ego(g, 2, nodes = which(grepl("Genotype", V(g)$type)), mode = "all") %>%
  map(function(.x) .x[grepl("MLG", names(.x))])
lengths(in_reach) %>%
  enframe(name = "index", value = "N") %>%
  ggplot(aes(x = N - 1)) +
  geom_histogram(binwidth = 1, center = 0, color = "black", fill = "grey50") +
  theme_bw(base_size = 16, base_family = "Helvetica") +
  theme(aspect.ratio = 1/2) +
  theme(axis.text = element_text(color = "black")) +
  scale_y_continuous(expand = c(0, 0.1)) +
  scale_x_continuous(expand = c(0, 0.1), breaks = c((0:6)*10), limits = c(-1, 60)) +
  theme(panel.grid.major = element_line(colour = "grey20")) +
  theme(panel.grid.minor = element_line(colour = "grey50", linetype = 3)) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(panel.border = element_blank()) +
  xlab("Number of MLG Neigbors")
```

The maximum here is `r max(lengths(in_reach))`, which represents MLG 25.

### Interactive Visualizations

@@

```{r top5graph, fig.height = 7, fig.width = 7}
top5 <- filter(mcgmlg, as.character(MCG) %in% mcgs$MCG[1:5])
top5g <- make_mcgmlg_graph(top5)
tosize <- V(top5g)$size
V(top5g)$size <- sqrt(tosize)/10
set.seed(2017-05-08)
top5lay <- create_layout(top5g, layout = "igraph", algorithm = "nicely")
top5lay$community <- inner_join(data_frame(vertex = top5lay$name), the_communities)$comm
t5g <- mcg_mlg_graph %+% top5lay + ggtitle("Top 5 Mycelial Compatibility Groups and associated MLGs")
t5g
ggsave(t5g, file = file.path("results", "figures", "publication", "FigureS2B.pdf"), width = 88, height = 88, scale = 3, units = "mm")
```

@@

severity_plot <- ggplot(Severity, aes(x = MCG, y = Severity)) +
  geom_point(aes(fill = MLG), 
             position = position_jitter(width = 0.2),
             alpha = 0.75,
             pch = 21) +
  scale_fill_viridis(discrete = TRUE, direction = -1) +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(aspect.ratio = 0.6) +
  facet_wrap(~Region, nrow = 2) +
  ylim(c(3.5, 8)) +
  labs(list(
    title = "Severity by MCG and Region",
    fill = "Multilocus Genotype",
    subtitle = "Five most abundant multilocus genotypes shown"
  ))
severity_plot
ggsave(severity_plot, filename = "results/figures/publication/FigureS3.pdf", width = 183, unit = "mm")
```

## Assessing MCG/MLG community structure/cohesiveness

>>>>

# MLG table

@@

# Making a Graph

@@

```{r community_counting, fig.width = 6, fig.height = 3.5}
in_reach <- ego(g, 2, nodes = which(grepl("Genotype", V(g)$type)), mode = "all") %>%
  map_df(function(.x) as_tibble(list(MLG = list(names(.x[grepl("MLG", names(.x))])), 
                                     MCG = list(names(.x[!grepl("MLG", names(.x))])), 
                                     WHO = names(.x[1])
                                     )
                                )) %>% 
  mutate(N = lengths(MLG))

in_reach_summary <- in_reach %>% 
  group_by(N) %>%
  summarize(E = diversity_stats(as.integer(table(unlist(MCG))))["E.5"], 
            MCG = n_distinct(unlist(MCG)), 
            MLG = n())
knitr::kable(in_reach_summary)
ggplot(in_reach_summary, aes(x = N - 1, y = MLG, fill = E)) +
  geom_segment(aes(yend = 0, xend = N-1)) +
  geom_point(aes(size = MCG), pch = 21) +
  # ggrepel::geom_label_repel(aes(label = MCG)) +
  viridis::scale_fill_viridis(option = "D") +
  theme_bw(base_size = 16, base_family = "Helvetica") +
  theme(aspect.ratio = 1/1.75) +
  theme(axis.text = element_text(color = "black")) +
  scale_y_continuous(expand = c(0, 0.1), limits = c(0, 25)) +
  scale_x_continuous(expand = c(0, 0.1), breaks = c((0:6)*10), limits = c(-2, 60)) +
  theme(panel.grid.major = element_line(colour = "grey20")) +
  theme(panel.grid.minor = element_line(colour = "grey50", linetype = 3)) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(panel.border = element_blank()) +
  labs(list(fill = expression(E[5]), 
            size = "MCG"
            # caption = "size and fill represent\nthe number and distribution\nof MCG"
            )) +
  xlab("Maximum Heterothallic Pairings") +
  ylab("Multilocus Genotypes")
```

The maximum here is `r in_reach %>% arrange(N) %>% tail(1) %>% pull(N)`, which 
represents `r in_reach %>% arrange(N) %>% tail(1) %>% pull(WHO)`.

## Interactive Visualizations

@@

```{r top5graph, fig.height = 7, fig.width = 7}
top5 <- filter(mcgmlg, as.character(MCG) %in% mcgs$MCG[1:5])
top5g <- make_mcgmlg_graph(top5)
tosize <- V(top5g)$size
V(top5g)$size <- sqrt(tosize)/10
set.seed(2017-05-08)
top5lay <- create_layout(top5g, layout = "igraph", algorithm = "nicely")
top5lay$community <- inner_join(data_frame(vertex = top5lay$name), the_communities)$comm
t5g <- mcg_mlg_graph %+% top5lay + ggtitle("Top 5 Mycelial Compatibility Groups and associated MLGs")
t5g
if (!interactive())
  ggsave(t5g, file = file.path("results", "figures", "publication", "FigureS2B.pdf"), width = 88, height = 88, scale = 3, units = "mm")
```

@@

severity_plot <- ggplot(Severity, aes(x = MCG, y = Severity)) +
  geom_point(aes(fill = MLG), 
             position = position_jitter(width = 0.2),
             alpha = 0.75,
             pch = 21) +
  scale_fill_viridis(discrete = TRUE, direction = -1) +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(aspect.ratio = 0.6) +
  facet_wrap(~Region, nrow = 2) +
  ylim(c(3.5, 8)) +
  labs(list(
    title = "Severity by MCG and Region",
    fill = "Multilocus Genotype",
    subtitle = "Five most abundant multilocus genotypes shown"
  ))
severity_plot
if (!interactive())
  ggsave(severity_plot, filename = "results/figures/publication/FigureS3.pdf", width = 183, unit = "mm")
```

# MCG accumulation permutations

On 2017-07-27, Sydney came up with the idea to assess MCG accumulation with 
increasing genetic distance similar to the genotype accumulation curve. First,
let's see what the curve would look like if we had one MLG/MCG.

First, we need to set up everything to calculate this quantity

```{r mlg accumulation calculation}
bd         <- bruvo.dist(dat11, replen = other(dat11)$REPLEN)
thresholds <- filter_stats(dat11, distance = bd, plot = TRUE)
thresh     <- c(0, unique(thresholds$farthest$THRESHOLDS) + .Machine$double.eps)
collapsed  <- map(thresh, . %>% mlg.filter(dat11, threshold = ., distance = bd))

# This function will count the number of original MLGs within each collapsed MLG
# Note that "original" here can be any MLG source.
compare_mlg <- function(collapsed, thresh, original){
  as_tibble(list(MLG = collapsed, y = original, threshold = thresh)) %>%
    group_by(MLG, threshold) %>%
    summarize(N = n_distinct(y)) %>%
    ungroup() %>%
    mutate(NMLG = n())
}

# A custom function for stat summary that returns a single point when there is
# only a single observation
my_meanse <- function (x, mult = 1) {
  x <- stats::na.omit(x)
  se <- mult * sqrt(stats::var(x)/length(x))
  mean <- mean(x)
  if (length(x) == 1) return(data.frame(y = mean, ymin = mean, ymax = mean))
  data.frame(y = mean, ymin = mean - se, ymax = mean + se)
}
```

Now, as a demontration, we can see what the curve would look like if all MCGs
were randmly distributed (e.g. not clumpy) among the MLGs

```{r mlg accumulation noclump, fig.width = 5, fig.height = 3}
map2_df(collapsed, thresh, compare_mlg, original = sample(87, 366, replace = TRUE)) %>%
  ggplot(aes(x = threshold, y = N)) +
  geom_point(alpha = 0.1) +
  stat_smooth(se = FALSE, color = "black") +
  stat_summary(pch = 21, fun.data = my_meanse, fill = "white")
```

Now we can see what it looks like when there is a 1:1 relationship of MCGs to
MLGs:

```{r mlg accumulation mlg, fig.width = 5, fig.height = 3}
map2_df(collapsed, thresh, compare_mlg, original = mll(dat11, "original")) %>%
  ggplot(aes(x = threshold, y = N)) +
  geom_point(alpha = 0.1) +
  stat_smooth(se = FALSE, color = "black") +
  stat_summary(pch = 21, fun.data = my_meanse, fill = "white")
```

Yet another comparison would be if each sample were a unique MCG:

```{r mlg accumulation uniq, fig.width = 5, fig.height = 3}
map2_df(collapsed, thresh, compare_mlg, original = seq(366)) %>%
  ggplot(aes(x = threshold, y = N)) +
  geom_point(alpha = 0.1) +
  stat_smooth(se = FALSE, color = "black") +
  stat_summary(pch = 21, fun.data = my_meanse, fill = "white")
```

And it's worth examining what happens if there are only 10 MCGs:

```{r mlg accumulation ten, fig.width = 5, fig.height = 3}
map2_df(collapsed, thresh, compare_mlg, original = sample(10, 366, replace = TRUE)) %>%
  ggplot(aes(x = threshold, y = N)) +
  geom_point(alpha = 0.1) +
  stat_smooth(se = FALSE, color = "black") +
  stat_summary(pch = 21, fun.data = my_meanse, fill = "white")
```

And, of course, we should see what our observed data looks like:

```{r mlg accumulation observed, fig.width = 5, fig.height = 3}
observed <- map2_df(collapsed, thresh, compare_mlg, original = mll(dat11, "custom")) %>%
  mutate(sample_number = "observed") %>%
  mutate(source = "observed")

# Revisiting random dist
map2_df(collapsed, thresh, compare_mlg, original = sample(87, 366, replace = TRUE)) %>%
  ggplot(aes(x = threshold, y = N)) +
  geom_point(alpha = 0.1) +
  stat_smooth(se = FALSE, color = "black") +
  scale_y_continuous(breaks = c(25, 50, 75)) +
  stat_summary(pch = 21, fun.data = my_meanse, fill = "white")

observed %>%
  ggplot(aes(x = threshold, y = N)) +
  geom_point(alpha = 0.1) +
  stat_smooth(se = FALSE, color = "black") +
  scale_y_continuous(breaks = c(25, 50, 75)) +
  stat_summary(pch = 21, fun.data = my_meanse, fill = "white")
```

```{r randomizations, cache = TRUE}
set.seed(5000)
random   <- map_df(1:20, function(x) map2_df(collapsed, thresh, compare_mlg, original = sample(mll(dat11, "custom"))), .id = "sample_number") %>%
  mutate(sample_number = sprintf("%3d", as.integer(sample_number))) %>%
  mutate(source = "random")
```

```{r visualization, fig.width = 5, fig.height = 3}

ggplot(random, aes(x = threshold, y = N, group = sample_number, color = source, fill = source)) +
  # geom_point(alpha = 0.01, color = "grey50") +
  stat_summary(alpha = 0.125, fun.data = my_meanse) +
  # geom_point(alpha = 0.5, pch = 21, data = observed, fill = "white") +
  stat_smooth(se = FALSE, data = observed) +
  stat_summary(pch = 21, fun.data = my_meanse, data = observed) +
  scale_color_manual(values = c("black", "grey50")) +
  scale_fill_manual(values = c("white", "grey50")) +
  theme_bw(base_size = 16, base_family = "Helvetica") +
  theme(aspect.ratio = 1/2) +
  theme(axis.text = element_text(color = "black")) +
  theme(panel.grid.major = element_line(colour = "grey20")) +
  theme(panel.grid.minor = element_line(colour = "grey50", linetype = 3)) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.minor.x = element_blank()) +
  theme(panel.border = element_blank()) +
  theme(legend.position = "top") +
  ylab("Number of MCG/MLG") +
  xlab("Bruvo's Distance")
```

# Assessing MCG/MLG community structure/cohesiveness
